---
title: "Elementary Statistics <br>Orientation"
author: "MTH-161D Elementary Statistics | Spring 2025 | University of Portland"
date: "January 13, 2025"
output:
  ioslides_presentation:
    widescreen: true
    transition: faster
bibliography: ../../references.bib
csl: ../../mbe.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives

- **Communicate highlights of the Syllabus**
- **Activity: Introductions**

## Course Information

- **Number:** MTH-161
- **Title:** Elementary Statistics
- **Section:** D
- **Term:** Spring 2025
- **Credits:** 3
- **Instructor:** Dr. Quijano

### **Schedule**

Monday, Wednesday, Friday at 9:15 AM to 10:10 AM

### **Location**

Franz Hall 125

## About the Instructor

### Dr. [Alex John Quijano](alexjohnquijano.com){target="_blank"}

- **Office:** Buckley Center 279
- **Email:** quijano@up.edu
- **Education and Profession:** Applied Mathematics
- **Teaching:** Probability and statistics, data science, linear algebra, calculus, differential equations
- **Research:** Mathematical interdisciplinary research in data science, modeling of natural languages, social and biological systems
- **Hobbies and Interests:** Movies, anime, musicals, coffee 
- **Other Facts:** Two cats

## Modes of Communication

### **Teams**

We will be using Teams as the main real-time communication tool for general announcements, question-answering discussions, and direct messages.

> The Teams for this course is [MTH-161D-sp25](https://teams.microsoft.com/l/team/19%3AzLuVe5GlJJ3Y5Ob-I0vypRDPEQLigxvlqaGHR7GAXv41%40thread.tacv2/conversations?groupId=a47c5106-9aa6-4569-b98f-17f70304580a&tenantId=ea8f3949-231c-40b6-a33f-56873af96f87){target="_blank"}. 

### **Class Website**

The syllabus, tentative topics schedule, and other information are posted on the course website.

> You can access the course website at [mth-161d-sp25](https://upmathematics.github.io/mth-161d-sp25/){target="_blank"}.

*Note that the course website can be accessed in Teams.*

## Structure and Expectations

### **Lectures and Discussions**

  - Pre-reading
  - Short lecture
  - Worksheet activity
  - Group presentations or discussions

### **Expectations**

  - Worksheets
  - Homework
  - Assessments

## Assignments

### **Worksheets:** Wednesday and Friday (due in class)

The purpose of the worksheets is for in-class group work and activities.

### **Homework:** Every other Monday (due Friday)

The purpose of the homework is to provide practice with problems, and it will include tasks that involve R.

### **Assessments:** Every other Monday (due in class)

The purpose of the assessments is to evaluate your understanding of the learning objectives.

## Submission Guidelines

### **Moodle**

We will be using the course [Moodle](https://learning.up.edu){target="_blank"} page for submitting assignments and giving feedback.

> The Moodle for this course is [MTH-161D - sp25 - Elementary Statistics](https://learning.up.edu/course/view.php?id=40831){target="_blank"}. You are already listed in Moodle for this course using your UP account.

## Resources

### **Textbook**

This course uses the following textbook - which is free to access - for reading and practice.

```{r echo=FALSE}
refs <- bibtex::read.bib("../../references.bib")
```

  * `r capture.output(print(refs["diez2012openintro"]))`
  
*Note that the textbook can be accessed in Teams.*
  
## Tools

### **R**

We will be using R, as a computational tool, for some assignments. These tools are free and open-source. 

> We will use the [Posit Cloud](https://login.posit.cloud/login?redirect=%2F){target="_blank"} service, an online computing environment.

*We will discuss more about this tool next time.*

## Academic Support

### **Help Hours**

> Dr. Alex John Quijano
> 
> * 4:00 PM - 5:00 PM Mondays to Fridays, Buckley Center 279
> * [One-to-One, Buckley Center 279 or in Teams](https://calendly.com/alexjohnquijano/helphours){target="_blank"} <br> *Click on the link to sign-up for a 15-minute session.*

My walk-in help hours start week 2.

### **Math Resource Center**

The [Math Resource Center](https://www.up.edu/learningcommons/tutoring-services/math-resource-center/index.html){target="_blank"} offiers walk-in tutoring and appointment-based tutoring.

## Activity

### Introductions

On a piece of paper:

  - Write your name, year, and major.
  - Discuss your background experience in math and science.
  - Discuss your anxieties and fears about math.
  - Discuss your hobbies.
  - Write an encouraging message to yourself six weeks from now.

Pair with another student and share what you wrote.

Submit your work.
