---
header-includes: |
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhead[LO,L]{\textbf{MTH-161 D} | Spring 2025}
  \fancyhead[RO,R]{\textbf{F 1/24 Worksheet} | Pg. \thepage}
  \fancyfoot[CO,C]{}
  \usepackage{float}
  \usepackage{multirow}
output:
  bookdown::pdf_document2:
    toc: no
    number_section: no
urlcolor: blue
link-citations: yes
bibliography: ../../../references.bib
csl: ../../../mbe.csl
geometry: "left=0.50in,right=0.50in,top=0.75in,bottom=0.75in"
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**Name:** \textcolor{blue}{\textbf{Solution}}

**Collaborators:**

\noindent\rule{\textwidth}{1pt}
**Instructions:** 

> You must submit your worksheet individually by end-of-class or end-of-day. Your name must exist in your worksheet and the names of your collaborators.

Worksheets are marked mostly on completion, and partially on correctness. It will be marked either pass or fail, there will no detailed feedback on worksheets, and no opportunities for revisions and make-up.
\noindent\rule{\textwidth}{1pt}

## Identify Types of Variables & Compute Descriptive Statistics

### 1. Air pollution and birth outcomes, study components.

> Researchers collected data to examine the relationship between air pollutants and preterm births in Southern California. During the study air pollution levels were measured by air quality monitoring stations. Specifically, levels of carbon monoxide ($CO$) were recorded in parts per million, nitrogen dioxide and ozone in parts per hundred million, and coarse particulate matter ($PM_{10}$) in $\mu g/m^3$. Length of gestation data were collected on 143,196 births between the years 1989 and 1993, and air pollution exposure during gestation was calculated for each birth. The analysis suggested that increased ambient $PM_{10}$ and, to a lesser degree, $CO$ concentrations may be associated with the occurrence of preterm births.

a. Identify the main research question of the study.

    \textcolor{blue}{
    The research question is ``Is there an association between air pollution exposure and preterm births?''.
    }

\vfill

b. Who are the subjects in this study, and how many are included?

    \textcolor{blue}{
    The cases are 143,196 births in Southern California between 1989 and 1993.
    }

\vfill

c. What are the variables in the study? Identify each variable as numerical or categorical. If numerical, state whether the variable is discrete or continuous. If categorical, state whether the variable is ordinal.

    \textcolor{blue}{
    There are three variables:
    \begin{itemize}
      \item Length of Gestation, a continuous numerical variable.
      \item Measurement levels of polutants, a continuous numerical variable.
      \item Types of polutants ($CO$, $PM_{10}$, etc.), a nominal categorical variable.
    \end{itemize}
    }

\vfill

\newpage

### 2. Understanding variable types and computing descriptive statistics

You are tasked with analyzing and comparing four data sets. Follow the steps below to identify the variable types, calculate descriptive statistics, and explain your reasoning.

| Set | Data |
|:---|:---|
| X | $3, 5, 5, 5, 8, 11, 11, 11, 13$ |
| Y | $3, 5, 5, 5, 8, 11, 11, 11, 20$ |
| A | $cat,cat,cat,dog,dog,mouse,mouse$ |
| B | $dog,mouse,cat,cat,cat,dog,cat$ |

a. Identify the type of variable and its sub type for each set.

    \textcolor{blue}{
    \begin{itemize}
    \item Set $X$ is a discrete numerical variable.
    \item Set $Y$ is a discrete numereical variable, the same as $X$.
    \item Set $A$ is a nominal categorical variable.
    \item Set $B$ is a nominal categorical variable, the same as $A$.
    \end{itemize}
    }

\vfill

b. Calculate descriptive statistics. For numerical data, compute the mean and median. For categorical data, calculate the frequency and proportion for each category.

    \textcolor{blue}{
    \textbf{For numerical variables $X$ and $Y$:}
    \begin{itemize}
    \item Set $X$ is $3, 5, 5, 5, 8, 11, 11, 11, 13$ with $9$ data points.
        $$
        \begin{aligned}
        & \text{The mean of set } X = \frac{3+5+5+5+8+11+11+11+13}{9} = 8 \\
        & \text{Sorted $X$} \longrightarrow 3, 5, 5, 5, \mathbf{8}, 11, 11, 11, 13 \\
        & \text{The median of set } X = 8 
        \end{aligned}
        $$
    \item Set $Y$ is $3, 5, 5, 5, 8, 11, 11, 11, 20$ with $9$ data points.
        $$
        \begin{aligned}
        & \text{The mean of set } Y = \frac{3+5+5+5+8+11+11+11+20}{9} = 8.778 \\
        & \text{Sorted $Y$} \longrightarrow 3, 5, 5, 5, \mathbf{8}, 11, 11, 11, 20 \\
        & \text{The median of set } Y = 8 
        \end{aligned}
        $$
    \end{itemize}
    }
    
    \vfill

    \newpage

    \textcolor{blue}{
    \textbf{For categorical variables $A$ and $B$:}
    \begin{itemize}
    \item Set $A$ is $cat,cat,cat,dog,dog,mouse,mouse$ with $7$ cases. 
        $$
        \begin{aligned}
        & \text{The unique categories are } cat, dog, \text{ and } mouse. & \\
        & \text{The frequency of each category is } & \\
        & & cat,cat,cat \longrightarrow 3 \\
        & & dog,dog \longrightarrow 2 \\
        & & mouse,mouse \longrightarrow 2 \\
        & \text{The proportion of each category is } & \\
        & & cat \longrightarrow \frac{3}{7} = 0.429 \\
        & & dog \longrightarrow \frac{2}{7} = 0.286 \\
        & & mouse \longrightarrow \frac{2}{7} = 0.286 \\
        \end{aligned}
        $$
    \item Set $B$ is $dog,mouse,cat,cat,cat,dog,cat$ with $7$ cases.
        $$
        \begin{aligned}
        & \text{The unique categories are } cat, dog, \text{ and } mouse. & \\
        & \text{The frequency of each category is } & \\
        & & cat,cat,cat,cat \longrightarrow 4 \\
        & & dog,dog \longrightarrow 2 \\
        & & mouse \longrightarrow 1 \\
        & \text{The proportion of each category is } & \\
        & & cat \longrightarrow \frac{4}{7} = 0.571 \\
        & & dog \longrightarrow \frac{2}{7} = 0.286 \\
        & & mouse \longrightarrow \frac{1}{7} = 0.143 \\
        \end{aligned}
        $$
    \end{itemize}
    }

\vfill

c. Provide a short written explanation for your findings. Be sure to justify why the chosen descriptive statistics are appropriate for the given variable type.

    \textcolor{blue}{
    The mean and median are best used with numerical variables because they involve calculations based on numbers, which are inherently quantitative. On the other hand, frequency and proportion are suited for categorical variables because they describe how often each category occurs, which is inherently qualitative.
    }

\vfill

\newpage

## \textcolor{blue}{R Commands (Optional)}

### \textcolor{blue}{2b}

\textcolor{blue}{Computing the mean and median of set $X$}.
```{r echo=TRUE}
X <- c(3,5,5,5,8,11,11,11,13)
mean(X)
median(X)
```

\textcolor{blue}{Computing the mean and median of set $Y$}.
```{r echo=TRUE}
Y <- c(3,5,5,5,8,11,11,11,20)
mean(Y)
median(Y)
```

\textcolor{blue}{Computing the frequencies and proportions of set $A$}.
```{r echo=TRUE}
A <- c("cat","cat","cat","dog","dog","mouse","mouse")
table(A)
table(A)/length(A)
```

\textcolor{blue}{Computing the frequencies and proportions of set $B$}.
```{r echo=TRUE}
B <- c("dog","mouse","cat","cat","cat","dog","cat")
table(B)
table(B)/length(B)
```

<!-- \newpage

## **References**

::: {#refs}
:::

<br> -->
